---
title: "COVID-19 Canada Testing"
runtime: shiny
output: 
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    theme: cosmo
    includes:
      after_body: footer.html
      in_header: google-analytics.html
---

```{r setup, echo=FALSE,results='hide',message = FALSE, warning = FALSE}


# deployApp("C:/DB Mount/GitHub/Testing_Dash", appName = "Testing_Dash") #for Kuan only;
#------------------ Packages ------------------
library(shiny) #shiny app;
library(flexdashboard) #dashboard;
library(dplyr) #data tool;
library(plotly) #visual tool;
library(tibble) #scrap variable characters;
library(DT) #Datatable;

#------------------ Data is generated by running the data clean script ------------------
# We run data cleaning in a separate R file in order to speed up the app loading time
#------------------ Read all data ------------------

can<-read.csv("docs/can.csv", header = TRUE, sep = ",", encoding = 'UTF-8')
can$date_report<-as.Date(can$date_report,format="%Y-%m-%d")
prov<-read.csv("docs/prov.csv", header = TRUE, sep = ",", encoding = 'UTF-8')
prov$date_report<-as.Date(prov$date_report,format="%Y-%m-%d")

df_trajectory<-read.csv("docs/df_trajectory.csv", header = TRUE, sep = ",", encoding = 'UTF-8')
df_trajectory_can<-read.csv("docs/df_trajectory_can.csv", header = TRUE, sep = ",", encoding = 'UTF-8')

can_testinginfo<-read.csv("docs/can_testinginfo.csv", header = TRUE, sep = ",")
world_testinginfo<-read.csv("docs/world_testinginfo.csv", header = TRUE, sep = ",")

#------------------ creating New bubble map data;  ------------------
# instead of using polygon maps, we will display bubble map it reads faster;
# only information needed are the bubble centre lat and lng for each province and territory;
library(leaflet) #mapping;
canada_region<- prov[prov$date_report==max(prov$date_report)&prov$province != "Repatriated",]
canada_region$province<-factor(canada_region$province)
canada_region$lat<-c(53.9333, 49.2827, 53.7609, 46.5653, 53.1355, 44.6820, 70.2998, 64.8255, 51.2538, 46.5107, 52.9399, 52.9399, 64.2823)
canada_region$lng<-c(-116.5765, -123.1207,  -98.8139,  -66.4619,  -57.6604, -63.7443, -83.1076 ,-124.8457, -85.3232  ,-63.4168,  -73.5491, -106.4509, -135.0000)
# Prepare the text for the tooltip:
mytext <- paste(
   canada_region$province, "<br/>",
   "Confirmed Cases: ", canada_region$cumulative_cases, "<br/>", 
   "Active Cases: ", canada_region$active_cases, "<br/>", 
   "Deaths: ", canada_region$cumulative_deaths, "<br/>", 
   "Recovered: ", canada_region$cumulative_recovered, "<br/>", 
   "Total tested: ", canada_region$ccumulative_testing, "<br/>", 
   "Total tested per 1000 population: ", canada_region$cumulative_testing_per1000, sep="") %>%
  lapply(htmltools::HTML)

#------------------ data update time;  ------------------
date_update_date<-read.table("https://raw.githubusercontent.com/ishaberry/Covid19Canada/master/update_time.txt")
# date_update_date<-read.table("docs/data/update_time.txt")
data_date<-as_tibble(paste(date_update_date$V1, date_update_date$V2, date_update_date$V3))


#------------------ Parameters colours  ------------------
# Set colors
# https://www.w3.org/TR/css-color-3/#svg-color
#value box colour;
tested_color <- "#000000"
tested_color2 <- "#00008B"
positive_color <- "#DC143C"
active_color <- "#FFA500"
recovered_color <- "forestgreen"
death_color <- "#FF0000"
ab_color<-"#000000"
bc_color<-"#E69F00"
mb_color<-"#800000"
nb_color<-"#CC79A7"
nl_color<-"#56B4E9"
ns_color<-"#808000"
nt_color<-"#625D5D"
nu_color<-"#1589FF"
on_color<-"#009E73"
pe_color<-"#0000A0"
qc_color<-"#0072B2"
sk_color<-"#800080"
yt_color<-"#E56717"


```


Summary {data-orientation=rows}
=======================================================================

Row
-----------------------------------------------------------------------
### tested {.value-box}
```{r}
valueBox(value = paste(format(max(can$cumulative_testing, na.rm = T), big.mark = ","), "", sep = " "), 
         caption = "Total Tested", 
         icon = "fas fa-user-md", 
         color = tested_color)
```


### tested2 {.value-box}
```{r}
valueBox(value = paste(format(max(can$cumulative_testing_per1000, na.rm = T), big.mark = ","), "", sep = " "), 
         caption = "Total Tested per 1000", 
         icon = "fas fa-user-md", 
         color = tested_color2)
```


### Positive Cases {.value-box}

```{r}
valueBox(value = paste(format(max(can$cumulative_cases, na.rm = T), big.mark = ","), "", sep = " "), 
         caption = "Total Confirm Cases", 
         icon = "far fa-plus-square", 
         color = positive_color)
```


### Active {.value-box}
```{r}
activecase<-max(can$cumulative_cases, na.rm = T) - max(can$cumulative_recovered, na.rm = T) - max(can$cumulative_deaths, na.rm = T)
valueBox(value = paste(format(activecase, big.mark = ","), sep = ""),
         caption = "Active Cases", 
         icon = "fas fa-ambulance",
         color = active_color)
```

### recovered {.value-box}
```{r}
valueBox(value = paste(format(max(can$cumulative_recovered, na.rm = T), big.mark = ","), sep = ""),
         caption = "Recovered Cases", 
         icon = "fas fa-heartbeat", 
         color = recovered_color)
```

### death {.value-box}
```{r}
valueBox(value = paste(format(max(can$cumulative_deaths, na.rm = T), big.mark = ","), sep = ""),
         caption = "Deaths", 
         icon = "", 
         color = death_color)
```



Row 
-----------------------------------------------------------------------

### Geographic Mapping on Cumulative Testing per 1000 Population in Canada

```{r}

# Create a color palette with handmade bins.
mybins <- c(0,100,150,200, 250,300,350,Inf)
pal <- colorBin( palette="viridis", domain=canada_region$cumulative_testing_per1000, na.color="transparent", bins=mybins)

renderLeaflet({leaflet(canada_region) %>%
  addTiles() %>%
  setView(lat=55, lng=-100, zoom = 3) %>%
    addCircleMarkers(~lng, ~lat, fillColor = ~pal(cumulative_testing_per1000), fillOpacity = 0.7, color = "red",radius = ~sqrt(cumulative_testing_per1000)*2, stroke = T, weight=2,label = mytext, labelOptions = labelOptions( style = list("font-weight" = "normal", padding = "3px 8px"), textsize = "12px", direction = "auto")) %>% addLegend( pal=pal, values=~cumulative_testing_per1000, opacity=0.9, title = "Cumulative Testing per 1000", position = "topright")})


```


### Confirmed Cases per 1000 Population by Total Tested per 1000 Population as of `r data_date$value`

```{r}
teamcolour<-c("#000000","#E69F00","#800000","#CC79A7","#56B4E9","#808000","#625D5D","#1589FF","#009E73","#0000A0","#0072B2","#800080","#E56717")

canada_positive<-round(sum(canada_region$cumulative_cases)/sum(canada_region$cumulative_testing)*100,1)
canada_region$cumulative_cases_per1000<- (round(canada_region$cumulative_cases/canada_region$population, 3))*1000
x1v<-max(canada_region$cumulative_testing_per1000)+25
y1v<-(x1v)*(sum(canada_region$cumulative_cases)/sum(canada_region$cumulative_testing))

plot_ly(data = canada_region, x = ~cumulative_testing_per1000, y = ~cumulative_cases_per1000, 
        color = ~province, colors = teamcolour, marker = list(size = 10)) %>% 
  layout( shapes=list(type='line', x0= 0, x1= x1v, y0=0, y1=y1v, line=list(dash='dot', width=2,color="Red")),
    title=paste("Nationwide Test-Case Ratio at ",canada_positive,"%",sep = ""),
    xaxis = list(title="Cumulative Testing per 1000 Population"),
    yaxis = list(title="Cumulative Cases per 1000 Population"))



```



Row 
-----------------------------------------------------------------------

### Canada COVID-19 Data by Province and Territory as of `r data_date$value`

```{r}


DT::renderDT(canada_region[,c(1,13,3,4,7,9,6,5,10,11,14)] %>%
  DT::datatable(rownames = FALSE,
                colnames = c("Region","Population","Daily New Cases","Cumulative Cases","Active Cases",
                             "Daily New Deaths", "Cumulative Deaths", "Cumulative Recovered", 
                             "Daily New Testing",
                             "Cumulative Testing","Cumulative Testing per 1000 Population"), options = list(order = list(1, 'desc'), searching = FALSE, pageLength = 13), filter = 'none'))
```



Trend
=======================================================================


Row
-----------------------------------------------------------------------

### Cumulative Number of Testing Completed in Canada

```{r}

plot_ly(prov, x = ~date_report, y = ~cumulative_testing, color = ~province, 
        colors = teamcolour, connectgaps = TRUE,  type = 'scatter',mode='lines') %>% 
  layout( xaxis = list(title = "",showgrid = FALSE),
         yaxis = list(title = "Cumulative Testing",
         showgrid = FALSE),hovermode = "compared")

```

### Cumulative Number of Testing Completed per 1000 Population in Canada

```{r}

plot_ly(prov, x = ~date_report, y = ~cumulative_testing_per1000, color = ~province, 
        colors = teamcolour, connectgaps = TRUE,  type = 'scatter',mode='lines') %>% 
  layout( xaxis = list(title = "",showgrid = FALSE),
         yaxis = list(title = "Cumulative Testing per 1000 Population",
         showgrid = FALSE),hovermode = "compared")

```

Row
-----------------------------------------------------------------------

### Estimated Daily Number of Testing Completed in Canada


```{r}
prov$testing[prov$testing <= 0]<-NA
plot_ly(prov, x = ~date_report, y = ~testing, color = ~province, 
        colors = teamcolour, connectgaps = TRUE,  type = 'scatter',mode='lines') %>% 
  layout( xaxis = list(title = "",showgrid = FALSE),
         yaxis = list(title = "Estimated Daily Testing",
         showgrid = FALSE),hovermode = "compared")

```


### Estimated Daily Percentage of Positive Testings


```{r}
prov2<- prov[prov$date_report>as.Date("2020-03-15"),]
prov2$positve<-round(prov2$cases/prov2$testing,2)*100
prov2$positve[prov2$positve<0]<-NA
prov2$positve[prov2$testing==0]<-NA
prov2$positve[prov2$positve>=100]<-NA

plot_ly(prov2, x = ~date_report, y = ~positve, color = ~province, 
        colors = teamcolour, connectgaps = TRUE,  type = 'scatter',mode='lines') %>% 
  layout( xaxis = list(title = "",showgrid = FALSE),
         yaxis = list(title = "Estimated Daily Percentage of Positives in %",
         showgrid = FALSE,ticksuffix = '%'),hovermode = "compared")

```


Trajectory
=======================================================================

Row
-----------------------------------------------------------------------

### Trajectory of Cumulative Testing in Canada Comparing to Other Countries

```{r}

plotly::plot_ly(data = df_trajectory) %>%
  plotly::add_lines(x = ~ index,
                    y = ~ can1,
                    name = "Canada", connectgaps = TRUE, line = list(width = 4,color = "red")) %>%
  plotly::add_lines(x = ~ index,
                    y = ~ us1,
                    line = list(color = "black", width = 2),
                    name = "United States",connectgaps = TRUE) %>%
   plotly::add_lines(x = ~ index,
                    y = ~ it1,
                    line = list(color = "#ec008c", width = 2),
                    name = "Italy",connectgaps = TRUE) %>%
   plotly::add_lines(x = ~ index,
                    y = ~ uk1,
                    name = "United Kingdom", connectgaps = TRUE, line = list(width = 2,color = "green")) %>%
   plotly::add_lines(x = ~ index,
                    y = ~ sko1,
                    line = list(color = "#00bcf2", width = 2),
                    name = "South Korea",connectgaps = TRUE) %>%
   plotly::add_lines(x = ~ index,
                    y = ~ au1,
                    line = list(color = "brown", width = 2),
                    name = "Australia",connectgaps = TRUE) %>%
   plotly::layout(yaxis = list(title = "Cumulative Testing (log-scale)",type = "log"),
                 xaxis = list(title = "Days since the total deaths surpass 100"))

```



### Trajectory of Cumulative Testing per 1000 Population in Canada Comparing to Other Countries

```{r}

plotly::plot_ly(data = df_trajectory) %>%
  plotly::add_lines(x = ~ index,
                    y = ~ can2,
                    name = "Canada",  line = list(width = 4,color = "red"),connectgaps = TRUE) %>%
  plotly::add_lines(x = ~ index,
                    y = ~ us2,
                    line = list(color = "black", width = 2),
                    name = "United States",connectgaps = TRUE) %>%
   plotly::add_lines(x = ~ index,
                    y = ~ it2,
                    line = list(color = "#ec008c", width = 2),
                    name = "Italy",connectgaps = TRUE) %>%
   plotly::add_lines(x = ~ index,
                    y = ~ uk2,
                    name = "United Kingdom",  connectgaps = TRUE,line = list(width = 2,color = "green")) %>%
   plotly::add_lines(x = ~ index,
                    y = ~ sko2,
                    line = list(color = "#00bcf2", width = 2),
                    name = "South Korea",connectgaps = TRUE) %>%
   plotly::add_lines(x = ~ index,
                    y = ~ au2,
                    line = list(color = "brown", width = 2),
                    name = "Australia",connectgaps = TRUE) %>%
   plotly::layout(yaxis = list(title = "Cumulative Testing per 1000 Population (log-scale)",type = "log"),
                 xaxis = list(title = "Days since the total deaths surpass 100"))

```


Row
-----------------------------------------------------------------------

### Trajectory of Cumulative Testing by Province

```{r}
plotly::plot_ly(data = df_trajectory_can) %>%
  plotly::add_lines(x = ~ index,
                    y = ~ can1,
                    name = "Canada", connectgaps = TRUE, line = list(width = 4,color = "red")) %>%
  plotly::add_lines(x = ~ index,
                    y = ~ qc1,
                    name = "Quebec", connectgaps = TRUE, line = list(width = 2, color=qc_color)) %>%
  plotly::add_lines(x = ~ index,
                    y = ~ on1,
                    name = "Ontario",connectgaps = TRUE, line = list(width = 2, color=on_color)) %>%
  plotly::add_lines(x = ~ index,
                    y = ~ ab1,
                    name = "Alberta", connectgaps = TRUE, line = list(width = 2, color=ab_color)) %>%
  plotly::add_lines(x = ~ index,
                    y = ~ bc1,
                    name = "British Columbia", connectgaps = TRUE, line = list(width = 2, color=bc_color))%>%
  plotly::add_lines(x = ~ index,
                    y = ~ ns1,
                    name = "Nova Scotia", connectgaps = TRUE, line = list(width = 2,color = ns_color))%>%
   plotly::layout(yaxis = list(title = "Cumulative Testing (log-scale)",type = "log"),
                 xaxis = list(title = "Days since the total deaths surpass 50"))


```

### Trajectory of Cumulative Testing per 1000 Population by Province

```{r}
plotly::plot_ly(data = df_trajectory_can) %>%
  plotly::add_lines(x = ~ index,
                    y = ~ can2,
                    name = "Canada", connectgaps = TRUE, line = list(width = 4,color = "red")) %>%
  plotly::add_lines(x = ~ index,
                    y = ~ qc2,
                    name = "Quebec", connectgaps = TRUE, line = list(width = 2, color=qc_color)) %>%
  plotly::add_lines(x = ~ index,
                    y = ~ on2,
                    name = "Ontario",connectgaps = TRUE, line = list(width = 2, color=on_color)) %>%
  plotly::add_lines(x = ~ index,
                    y = ~ ab2,
                    name = "Alberta", connectgaps = TRUE, line = list(width = 2, color=ab_color)) %>%
  plotly::add_lines(x = ~ index,
                    y = ~ bc2,
                    name = "British Columbia", connectgaps = TRUE, line = list(width = 2, color=bc_color))%>%
  plotly::add_lines(x = ~ index,
                    y = ~ ns2,
                    name = "Nova Scotia", connectgaps = TRUE, line = list(width = 2,color = ns_color))%>%
   plotly::layout(yaxis = list(title = "Cumulative Testing (log-scale)",type = "log"),
                 xaxis = list(title = "Days since the total deaths surpass 50"))

```




Canada Data
=======================================================================

```{r}
DT::renderDT(can[,c(1,4,7,5,6,11,14,12)] %>% 
  DT::datatable(rownames = FALSE,
                colnames = c("Date","Cumulative Cases","Active Cases", "Cumulative Recovered", "Cumulative Death", "Cumulative Testing", "Cumulative Testing per 1000","Testing Info"), options = list(searchHighlight = TRUE,  pageLength = 100), filter = 'top'))
```




About
=======================================================================


**Time of last data update:** `r data_date$value`



**1. References**

  - Berry I, Soucy J-PR, Tuite A, Fisman D. Open access epidemiologic data and an interactive dashboard to monitor the COVID-19 outbreak in Canada. CMAJ. 2020 Apr 14;192(15):E420. doi: https://doi.org/10.1503/cmaj.75262
  - World data used in this dashboard are from Our World In Data, https://github.com/owid/covid-19-data/tree/master/public/data/.
    
**2. Project Members**

  - Project Lead, [Kuan Liu](https://twitter.com/KuanLiu2), PhD Candidate in Biostatistics, Dalla Lana School of Public Health, University of Toronto
  - Data Manager, [Thai-Son Tang](https://twitter.com/ThaiSonTang), PhD student in Biostatistics, Dalla Lana School of Public Health, University of Toronto
  - [Rose Garrett](https://twitter.com/rose_carrot), PhD student in Biostatistics, Dalla Lana School of Public Health, University of Toronto 
  - [Alex Bushby](https://twitter.com/bushby_alex), PhD student in Biostatistics, Dalla Lana School of Public Health, University of Toronto
  - [Maxwell Garrett](https://github.com/maxwell-garrett), Undergraduate student - Computer Science Specialist, Faculty of Arts & Science, University of Toronto
  - Project Mentors
    - Dr. Eleanor Pullenayegum, Senior Scientist, Child Health Evaluative Sciences, The Hospital for Sick Children
    - Dr. Anna Heath, Scientist, Child Health Evaluative Sciences, The Hospital for Sick Children
 


**3. Feedbacks**

You can find the code of this dashboard at <https://github.com/Kuan-Liu/Testing_Dash>. Please feel free to reach me at kuan.liu@mail.utoronto.ca. You can also find me on [Twitter](https://twitter.com/KuanLiu2).


Row
-----------------------------------------------------------------------

### A. Testing Unit Information from the COVID-19 Canada Open Data Working Group 	[(Link to Data)](https://github.com/ishaberry/Covid19Canada)


```{r}
names(can_testinginfo)<-c("Provinces and Territories", "Testing Unit", "Testing Info")
knitr::kable(can_testinginfo)
```

### B. Testing Unit Information from the Our World in Data COVID-19 data	[(Link to Data)](https://github.com/owid/covid-19-data/tree/master/public/data/)


```{r}
names(world_testinginfo)<-c("Countires", "Testing Unit")
knitr::kable(world_testinginfo)
```

